---
title: "International Consensus: A Look at the Voting History of the United Nations General Assembly"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# set T for report.md, F for readme.md
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(unvotes)
library(lubridate)
library(broom)
library(tidyr)
library(purrr)
library(knitr)
```

## Introduction

First convened in January 1946, the United Nations General Assembly (UNGA) functions as the primary forum for crafting policy in the United Nations. 51 countries were represented at the inaugural meeting; today (2023) the body comprises 193 member states.

Over the almost eight decades of its existence, the UNGA has deliberated resolutions concerning a range of topics, from human rights to nuclear weapons to economic development. With data scientist David Robinson's R package `unvotes`, one can gain a better understanding of not only how the member states have voted over time, but also which issues have—or have not—garnered international consensus.  

### Data Preprocessing

We can see that there are three datasets in the `unvotes` package: `un_votes`, `un_roll_calls`, and `un_roll_call_issues`.

The first, `un_votes`, contains data regarding each member state's name (`country`), country code (`country_code`, such as US for the United States), roll call ID (`rcid`), and vote (`vote`) for a given resolution. Votes are categorized as "yes", "no", or "abstain". (States that were absent during the vote are not included in this dataset.) 

```{r}
head(un_votes)
```

The second dataset, `un_roll_calls`, describes each roll call vote, including the date (`date`), the resolution being voted on (`unres`), and a description of said resolution (`descr`). There are also variables regarding the session number (`session`, with 1 representing 1946), whether the vote was deemed "important" by the U.S. State Department (`importantvote`, being either no/0 or yes/1) and if the vote was on an amendment to a resolution (`amend`) or a paragraph in a resolution (`para`).

It is important to note that the data spans from the first session (1946) to the seventy-third session (2019).

```{r}
head(un_roll_calls)
```

The final dataset, `un_roll_call_issues`, details the issue to which each vote pertained. There are a total of six topics for resolutions in this dataset, namely arms control and disarmament (abbreviated as "di"), the Israeli-Palestinian conflict ("me"), human rights ("hr"), colonialism ("co"), nuclear weapons and nuclear material ("nu"), and economic development ("ec").

```{r}
head(un_roll_call_issues)
```

The three aforementioned datasets are more or less "clean," but I wanted to join `un_votes` and `un_roll_calls` in order to understand more about each country's vote. Using `inner_join()` from `dplyr` on the roll call ID, I created a new tibble called `joined_votes`.

```{r}
# Join votes and roll call datasets
joined_votes <- un_votes %>%
  inner_join(un_roll_calls, by = "rcid")
head(joined_votes)
```

Furthermore, I re-coded the `vote` column for easier analysis, designating "no" as 0, "yes" as 1, and "abstain" as 3.

```{r}
# Recode 'vote' column as no = 0, yes = 1, abstain = 2
joined_votes <- joined_votes %>%
  mutate(vote = recode(vote,
                       no = 0,
                       yes = 1,
                       abstain = 2))
head(joined_votes)
```

Aside from those actions, I chose not to eliminate unnecessary columns or perform other preprocessing steps at the outset. I believed that it would be better to combine and streamline the datasets as I conducted my analysis, performing any "tidying" (mostly for aesthetic purposes) later on.

## Exploring the Data

### General Findings

We can begin the examination of the data by calculating the total number of votes from 1946 to 2019 and the percentage of which were "yes".

```{r}
# Calculate total and fraction of 'yes' votes
total_plus_yes <- joined_votes %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# total_plus_yes[[1]] # total number of votes cast
# total_plus_yes[[2]] # percentage of yes votes
```

The total number of votes cast is `r total_plus_yes[[1]]`, and approximately `r round(total_plus_yes[[2]],4)*100`% of those votes were "yes". The latter value is not very meaningful, however, as it does not describe when, by whom, or about which issue the "yes" votes were made.

So, let us summarize the number of total votes and "yes" votes by year.

```{r}
library(lubridate)

# Summarize by year
joined_votes %>%
  group_by(year = year(date)) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))
```

We can use the data to make a line graph with the `ggplot2` package.

```{r}
# Summarize by year
by_year2 <- joined_votes %>%
  group_by(year = year(date)) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# Create line plot
ggplot(by_year2, aes(x=year, y=percent_yes)) +
  geom_line() + 
  labs(x="Year", y="Percentage of Yes Votes",
       title="Overall Voting by Year in UNGA") +
  scale_x_continuous(breaks = seq(1940, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# Peek at specific "dip" years (1948, 1960)
# by_year2 %>% filter(year == 1948)
# by_year2 %>% filter(year == 1960)
```

From the graph, it would seem that the consensus of U.N. member states has generally risen over time, although there appear to have been sharp disagreements at certain points. Indeed, the line "dips" around 1948, for example, which was perhaps related to the creation of the State of Israel and ensuing conflict. (Interestingly, the most contentious vote that year seemed to have been on a resolution to adopt Spanish as one of the working languages of the UNGA [A/RES/247(III)], with 32 members in favor, 20 against, and 5 abstaining). Another "dip" occurs around 1960, which could have been due to strong opposition to resolutions about colonial territories (e.g., A/RES/1580(XV)) and the devolving situation in the Congo (e.g., A/RES/1498(XV)). After the mid-1960's, `percent_yes` begins to rise relatively rapidly to 87% in 1977 and hovered between 75% and 85% in the years that followed.

A scatter plot presents a more nuanced view of the voting trends, with an added smoothing curve.

```{r}
# Create scatter plot with smoothing curve
ggplot(by_year2, aes(year, percent_yes)) +
  geom_point() + geom_smooth() + 
  labs(x="Year", y="Percentage of Yes Votes",
       title="Overall Voting by Year in UNGA") +
  scale_x_continuous(breaks = seq(1940, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

One can also organize the data by country:

```{r}
# Summarize by country
by_country2 <- joined_votes %>%
  group_by(country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))
by_country2

# Sort in ascending order of percent_yes
by_country_asc <- by_country2 %>% arrange(percent_yes)

# Sort in descending order
by_country_desc <- by_country2 %>% arrange(desc(percent_yes))
```

We can group the data by year and country to see how positively each member state voted in a given year.

```{r}
# Group by year and country
by_year_country2 <- joined_votes %>%
  group_by(year = year(date), country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))
by_year_country2
```

Upon further inspection, we can see that the country with the lowest overall percentage of "yes" votes (`percent_yes`) is `r by_country_asc$country[1]`, which voted "yes" `r round(by_country_asc$percent_yes[1],4)*100` of the time. The country with the highest percentage of positive votes is `r by_country_desc$country[1]`, and it voted "yes" in approximately `r round(by_country_asc$percent_yes[1],4)*100`% of its `r by_country_desc$total[1]` votes.

That being said, `r by_country_asc$country[1]` has only voted `r by_country_asc$total[1]` times, so its `percent_yes` does not signify much. 

```{r}
# Filter out countries with fewer than 100 votes
by_country2 <- by_country2 %>%
  arrange(percent_yes) %>%
  filter(total > 99)
```

After filtering out countries that cast less than 100 votes, the country with the lowest `percent_yes` is the `r by_country2$country[1]` (`r round(by_country2$percent_yes[1],4)*100`% of `r by_country2$total[1]` votes).

### Country-Specific Trends

Having examined patterns among all of the UNGA member states, we should also study individual countries' attitudes over time. The United Kingdom is one of the founding members of the Assembly and the U.N. Security Council, and it featured prominently in debates regarding the Israeli-Palestinian conflict (having previously governed the region) and colonialism, among other topics. 

```{r}
# Only UK data
UK_by_year2 <- by_year_country2 %>%
  filter(country == "United Kingdom")
UK_by_year2
```

We can filter the country-specific, annual "yes" vote averages for only the United Kingdom, and the year-to-year differences are best presented in the form of a line graph:

```{r}
# Line plot of percent_yes over time for UK only
ggplot(UK_by_year2, aes(year, percent_yes)) +
  geom_line(aes(group=1), color = "red") + 
  labs(x="Year", y="Percentage of Yes Votes",
       title="United Kingdom - UNGA Votes") +
  scale_x_continuous(breaks = seq(1940, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

One can observe from the graph that, except for a few brief periods, the votes by the U.K. did not often average above 60% in favor of U.N. resolutions. There was a spike to 100% "yes" votes in 1951; but only 7 resolutions were put to a vote that session, and they were largely procedural in nature.

France is another important player in world politics and was similarly instrumental in founding the United Nations and its various organs. We can compare its voting history with that of the U.K.

```{r}
# Filter for the UK and France
UK_FR_by_year2 <- by_year_country2 %>%
  filter(country %in% c("United Kingdom", "France"))
# UK_FR_by_year2

# Plot
ggplot(UK_FR_by_year2, aes(year, percent_yes, color=country)) +
  geom_line() +
  labs(x="Year", y="Percentage of Yes Votes", color="Country",
       title="United Kingdom and France - UNGA Votes") +
  scale_color_manual(values = c("goldenrod", "red")) +
  scale_x_continuous(breaks = seq(1940, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

The trends of the two states is not dissimilar, although the lines diverge significantly (by a margin of about 45%) between 1960 and 1965. That being said, they voted more or less the same from the 1980's into the new century.

Let us add in another power player, the United States, and a non-Western country, such as India, and see how the voting patterns compare:

```{r}
# Comparing four countries (US, UK, France, India)
countries2 <- c("United States", "United Kingdom",
                "France", "India")
filtered_4_countries2 <- by_year_country2 %>%
  filter(country %in% countries2)
filtered_4_countries2

# Line plot of % yes in four countries
ggplot(filtered_4_countries2, aes(year, percent_yes, color=country)) +
  geom_line() +
  labs(x="Year", y="Percentage of Yes Votes", color="Country",
       title="Multinational Comparison - UNGA Votes") +
  scale_color_manual(values = c("goldenrod", "chartreuse3", "red", "dodgerblue4")) +
  scale_x_continuous(breaks = seq(1940, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```
Now the story begins to get a bit more interesting. After the 1980's, the U.S. voted "yes" much less than either the U.K. or France, while India maintained a significantly higher "yes" average than the other three countries throughout the entire period.

Finally, we can view the voting records for 6 countries—the U.S., the U.K., France, Japan, Brazil, and India—over the last seventy years.

```{r}
# Comparing 6 countries (US, UK, France, Japan, Brazil, India)
countries3 <- c("United States", "United Kingdom",
                "France", "Japan", "Brazil", "India")
filtered_6_countries2 <- by_year_country2 %>%
  filter(country %in% countries3)
filtered_6_countries2

# Line plot of % yes over time faceted by country
ggplot(filtered_6_countries2, aes(year, percent_yes, color=country)) +
  geom_line() +
  labs(x="Year", y="Percentage of Yes Votes") +
  scale_color_manual(values = c("seagreen2", "goldenrod", "chartreuse3", "orchid2", "red", "dodgerblue4")) +
  scale_x_continuous(breaks = seq(1940, 2020, by = 10)) +
  # scale_y_continuous(breaks = seq(0, 1, by = 0.5)) +
  theme(legend.position = "none") +
  facet_wrap(~ country)
# + facet_wrap(~ country, scales = "free_y") # for free y-axis
```

Japan appears to have been somewhat more stable than the Western countries, averaging between 50% and 75% "yes", broadly speaking. Brazil, like India, voted "yes" at a much higher rate than the Western members, particularly after 1970.

That being said, visualizing these trends can only tell us so much, and so we need to quantify these trends, with the help of linear regression. We can use the United States as an example.

```{r}
# Calculate percentage of yes votes from the US by year
US_by_year2 <- by_year_country2 %>%
  filter(country == "United States")
# US_by_year2

# Perform a linear regression of percent_yes by year
US_fit2 <- lm(percent_yes ~ year, US_by_year2)
# US_fit2

# Perform summary() on the US_fit2 object
summary(US_fit2)
```

Working with `percent_yes` as the dependent variable and `year` as the independent variable, our model shows that the slope is `r round(US_fit2$coefficients[2],4)`. Furthermore, the $p$-value is less than 0.05 ($a$), meaning that the relationship between `percent_yes` and `year` is statistically significant.

Likewise, we can run a linear regression model for the United Kingdom and compare the results to those from the U.S. model.

```{r}
library(broom)

# Fit model for the UK
UK_by_year2 <- by_year_country2 %>% filter(country=="United Kingdom")
# UK_by_year
UK_fit2 <- lm(percent_yes ~ year, UK_by_year2)
# UK_fit

# Create US_tidied and UK_tidied
US_tidied2 <- tidy(US_fit2)
# US_tidied2
UK_tidied2 <- tidy(UK_fit2)
# UK_tidied2

# Combine the two tidied models
US_UK_tidied <- bind_rows(US_tidied2, UK_tidied2)
US_UK_tidied
```

In relation to the U.S., the U.K. trend line has a slightly positive slope (`r round(US_UK_tidied$estimate[4],4)`). Although the U.K. $p$-value (`r round(US_UK_tidied$p.value[4],4)`) is greater than that of the U.S., it is still less than $a$ and therefore suggests a significant relationship between `percent_yes` and `year` for the U.K.

We can go one step further and run linear regression on each country in the UNGA, after nesting the data by country.

```{r}
library(tidyr)

# Nest all columns besides country
nested2 <- by_year_country2 %>% nest(-country)
nested2
# nested2$data[[1]] # Afghanistan's voting history (percent_yes)

library(purrr)

# Perform a linear regression on each item in the data column, add another
# mutate that applies tidy() to each model, unnest the tidied column
country_coefficients2 <- by_year_country2 %>%
  nest(-country) %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, data = .)),
         tidied = map(model, tidy)) %>%
  unnest(tidied)
country_coefficients2
```

Having both the intercept and slope terms makes this tibble somewhat unsightly, and so we can filter only for the slope terms (`year`).

```{r}
# Filter for only slope terms (year, i.e. not intercept), 
# perform p-value correction, filter for significant values
filtered_countries2 <- country_coefficients2 %>% 
  filter(term=="year") %>%
  mutate(p.adjusted = p.adjust(p.value)) %>%
  filter(p.adjusted < 0.05)
filtered_countries2
```

The resulting tibble is a bit easier to read. I additionally adjusted all $p$-values to ensure accuracy and filtered for only ones that are significant (< $a$).   

Now we can determine which countries' voting records increased or decreased in terms of `percent_yes` the most sharply.

```{r}
# Sort for the countries which increased most sharply
filtered_countries2 %>% arrange(desc(estimate))

# Sort for the countries which decreased most sharply
filtered_countries2 %>% arrange(estimate)
```

The Polynesian island of Tuvalu had the sharpest increase in `percent_yes` over time, whereas Bosnia & Herzegovina showed the sharpest downturn in `percent_yes`.

### Trends by Issue

Let us turn next to the specific issues being voted upon, using the `un_roll_call_issues` dataset.

```{r}
# Count unique issues being discussed in UN resolution votes
count(un_roll_call_issues, issue, sort=TRUE)
```

U.N. resolutions were most related to arms control and disarmament with 1,092 resolutions, followed by the Israeli-Palestinian conflict with 1,061. Economic development, on the other hand, was the subject of only 765 resolutions.

We can join together our previous working tibble with these labels so that we can better discern how countries voted on particular issues.

```{r}
# Join together joined_votes, un_roll_call_issues based on the 'rcid' column
votes_combined <- joined_votes %>%
  inner_join(un_roll_call_issues, by = "rcid")
votes_combined
```

Having spent a good deal of time exploring the data, I felt that at this point I could eliminate unneccessary columns and create, reorder, or rename a few others. 

```{r}
# Clean up votes_combined: create variable 'year'; drop variables 'importantvote', 
# 'country_code', 'amend', 'para', 'short', 'descr', 'date'; 
# rename variable 'short_name' as 'issue_abrv'; recode 'Palestinian conflict'
# as 'Israeli-Palestinian conflict'
clean_votes <- votes_combined %>% 
  mutate(year = session + 1945) %>%
  select(c(rcid, session, year, country, vote, unres, 
           issue_abrv = short_name, issue)) %>%
  mutate(issue = recode(issue, "Palestinian conflict" = "Israeli-Palestinian conflict"))
clean_votes
```

One of the added columns is `issue_abrv`, which contains an abbreviation of the topic at hand for a given resolution. For instance, we can filter for "nu" to show only votes that pertained to nuclear weapons and nuclear material.

```{r}
# Filter for votes related to nuclear weapons and nuclear material (nu)
clean_votes %>% filter(issue_abrv=="nu")
```

Moreover, we can filter for only the United States' votes on nuclear weapons and graph them over time.

```{r}
# Filter, then summarize by year for US votes on nuclear weapons/material ('nu')
US_nu_by_year2 <- clean_votes %>% 
  filter(country == "United States", issue_abrv == "nu") %>%
  group_by(year) %>%
  summarize(percent_yes = mean(vote == 1))
# US_nu_by_year2

# Graph the percent of US 'yes' votes over time on nuclear weapons/material
ggplot(US_nu_by_year2, aes(year, percent_yes)) +
  geom_line(aes(group=1), color = "dodgerblue4") + 
  labs(x="Year", y="Percentage of Yes Votes",
       title="United States - Votes on Nuclear Weapons/Material") +
  scale_x_continuous(breaks = seq(1940, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

The line appears fairly chaotic, although votes after 1975 were generally less favorable than during the previous decades.

We can group each country's votes by both year and issue.

```{r}
# Group each country's votes by year, issue
by_country_year_issue <- clean_votes %>%
  group_by(country, year, issue) %>%
  summarize(total = n(), percent_yes = mean(vote == 1)) %>%
  ungroup()
by_country_year_issue
```

There is not much sense in graphing every country's voting history on each issue, but we can examine individual member states, such as the U.S.

```{r}
# Filter by_country_year_issue for just the US
US_by_country_year_issue <- by_country_year_issue %>%
  filter(country == "United States")
# US_by_country_year_issue

# Plot % yes over time for the US, faceting by issue
ggplot(US_by_country_year_issue, aes(year, percent_yes)) + 
  geom_line(aes(group=1), color = "dodgerblue4") + 
  labs(x="Year", y="Percentage of Yes Votes",
       title="United States - UNGA Votes by Issue") +
  scale_x_continuous(breaks = seq(1940, 2020, by = 10)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  facet_wrap(~ issue)
```

It seems that, on the whole, the U.S. has voted "yes" less on all six issues after 1975. How does the United Kingdom, our partner in the "Special Relationship", fare?

```{r}
# Filter by_country_year_issue for just the US and the UK
US_UK_by_country_year_issue <- by_country_year_issue %>%
  filter(country %in% c("United Kingdom", "United States"))
# US_UK_by_country_year_issue2

# Plot % yes over time for the US and UK, faceting by issue
ggplot(US_UK_by_country_year_issue, aes(year, percent_yes, color=country)) + 
  geom_line() + 
  labs(x="Year", y="Percentage of Yes Votes", color="Country",
       title="U.S. and U.K. - UNGA Votes by Issue") +
  scale_color_manual(values = c("red", "dodgerblue4")) +
  scale_x_continuous(breaks = seq(1940, 2020, by = 10)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  facet_wrap(~ issue)
```

While the U.K. also voted "yes" less after the 1970's, it was nevertheless more positive than the U.S., especially regarding the Israeli-Palestinian conflict.

Just as we did for each country's overall percentage of "yes" votes, we can fit a linear regression model for each country's votes by issue over time.

```{r}
# Fit model on the by_country_year_issue dataset
country_issue_coefficients <- by_country_year_issue %>%
  nest(-country, -issue) %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, data = .)),
         tidied = map(model, tidy)) %>%
  unnest(tidied)
# country_issue_coefficients

# Filter to only include slope terms, adjust p-values, filter for significant values
country_issue_filtered <- country_issue_coefficients %>%
  filter(term=="year") %>%
  mutate(p.adjusted = p.adjust(p.value)) %>%
  filter(p.adjusted < 0.05)
country_issue_filtered
```

Armed with this information, we can answer questions such as: Which country showed the steepest downward trend in `percent_yes`?

```{r}
# Identify country with the steepest downward trend in percent_yes
country_issue_filtered %>% arrange(estimate) # Vanuatu, re: Isr-Pal conflict
```

With a slope of `r round(min(country_issue_filtered$estimate),4)`, `r country_issue_filtered$country[which(country_issue_filtered$estimate == min(country_issue_filtered$estimate))]` has the most drastic negative trend of the UNGA member states for `percent_yes`.

As before, we can graph Vanuatu's voting patterns by issue.

```{r}
# Filter data for only Vanuatu's votes
vanuatu_by_country_year_issue <- by_country_year_issue %>% filter(country == "Vanuatu")
# vanuatu_by_country_year_issue

# Plot of percentage 'yes' over time, faceted by issue
ggplot(vanuatu_by_country_year_issue, aes(year, percent_yes)) + 
  geom_line(aes(group=1), color = "seagreen4") + 
  labs(x="Year", y="Percentage of Yes Votes",
       title="Vanuatu - UNGA Votes by Issue") +
  scale_x_continuous(breaks = seq(1940, 2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.05)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  facet_wrap(~ issue)
```

Granted, the island nation has only been a member of the United Nations since 1981, but certain trends are clear from the graphs: On the one hand, it seems that Vanuatu has voted very positively during its tenure in the UNGA. On the other hand, its record is interrupted by dramatic dips, particularly after the year 2000 and regarding the Israeli-Palestinian conflict. Vanuatu recognized the State of Palestine in 1989 and established diplomatic relations that same year. However, in 2011, Vanuatu's representatives made what appears to be an about-face, opposing Palestine's UNESCO application along with fourteen other member states. In short, the country's attitudes on the issue are as chaotic as the graph suggests.

## Conclusion

The United Nations General Assembly is unique, in that it was created to provide an equal forum for all countries to present their views on major issues. Using David Robinson's `unvotes` datasets, we have examined when and how member states have voted on various issues and considered reasons as to why they might have voted the way that they did. These datasets have stories to tell, and those stories are wrapped up in the story of humanity over the past century.

In the future, I hope to revisit this project and experiment with how machine learning could be used to predict countries' voting patterns in the future.

## Credits

This project was inspired by David Robinson's DataCamp course entitled "Case Study: Exploratory Data Analysis in R," which used abridged (and slightly different) versions of the datasets in `unvotes`.
